Best Practices for Alchemical Free Energy Calculations: v\versionnumber

### 2019-05-08 paper/manuscript.tex
Tidying of Prerequesites and Scope, introduction and Step 0 and step 1

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% ARTICLE SETUP
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\title{Best Practices for Alchemical Free Energy Calculations: v\versionnumber}
\author[7]{Bryce K. Allen}
\author[1*]{John D. Chodera}
\author[2*]{Antonia S. J. S. Mey}
\author[2*]{Julien Michel}
\author[3*]{David L. Mobley}
\author[1]{Levi N. Naden}
\author[4*]{Samarjeet Prasad}
\author[5*]{Julia E. Rice}
\author[1,9]{Andrea Rizzi}
\author[6*]{Michael Shirts}
\author[10*]{Huafeng Xu}

@@

\begin{abstract}
%\todo[inline, color={green!20}]{ASJSM: @Volunteer write abstract}
%In your work, in this particular slot, please provide an abstract of no more than 250 words.
%Your abstract should explain the main contributions of your article, and should not contain any material that is not included in the main text.
%Please note that your abstract, plus the authorship material following it, must not extend beyond the title page or modifications to the LaTeX class will likely be needed.
Alchemical free energy calculations can be a useful tool for predicting free energy differences associated with the transfer of small molecules from one environment to another.
The hallmark of these methods is the use of modified potential energy functions to represent \emph{alchemical} intermediate states that cannot exist in chemistry; by analyzing simulation data collected from a series of bridging alchemical thermodynamic states, transfer free energies (or differences in transfer free energies) can be computed with orders of magnitude less simulation time than observing the process spontaneously. 
While these methods are highly flexible, care must be taken in avoiding common pitfalls to ensure that computed free energy differences can be robust and reproducible for the chosen forcefield, and that appropriate corrections are included to permit comparison with experimental data.
In this paper, we review current best practices for several popular application domains of alchemical free energy calculations, including relative and absolute small molecule binding free energy calculations to biomolecular targets.
\todo[inline]{JDC: What about biopolymer mutations?}
\end{abstract}

@@

%%%%%%%%%%%%%%%%%%%%
%              Introduction                  %
%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
\label{sec:intro}
Alchemical free energy calculations have become a mature technology for computing various properties related to the transfer of chemical species from one environment to another.
The domain of applicability for these calculations now involves such varied applications as the computation of protein-ligand binding free energies~\cite{binding-free-energies}, ligand selectivities~\cite{selectivity}, partition and distribution coefficients between different liquid phases (such as octanol-water partition coefficients~\cite{octanol-water-partition}), loss of affinity due to resistance mutations~\cite{hauser2018predicting,aldeghi2018accurate}, and changes in protein thermostability due to engineered mutations~\cite{seeliger2010protein,gapsys2016insights,gapsys2016accurate,aldeghi2019accurate}.

The defining characteristic of alchemical free energy calculations is the use of a series of alchemically-modified potential functions $U(x; \lambda)$ in which an alchemical parameter $\lambda$ modulates interactions in a manner that cannot occur in real chemical systems.
One or more simulations are used to collect data from a multitude of alchemical states to compute a free energy difference between a chemical state ($\lambda_0$) and another chemical or alchemical reference state ($\lambda_1$),
\begin{eqnarray}
\Delta f &\equiv& f(\lambda_1) - f(\lambda_0) = - \ln \frac{Z(\lambda_1)}{Z(\lambda_0)} , \label{equation:dimensionless-free-energy-difference}
\end{eqnarray}
where the dimensionless free energy $f(\lambda) \equiv \beta F(\lambda)$ is given in terms of partition functions $Z(\lambda)$,
\begin{eqnarray}
Z(\lambda) &=& \int dx \, e^{-u(x; \lambda)} . \label{equation:partition-function-definition}
\end{eqnarray}
Here, the inverse thermal energy $\beta \equiv (k_B T)^{-1}$ where $k_B$ is the Boltzmann constant and $T$ is the absolute temperature, and the \emph{reduced potential} $u(x; \lambda)$~\cite{shirts2008statistically} is generally given by a trace over thermodynamic parameters with their conjugate dynamical variables,
\begin{eqnarray}
u(x;\lambda) &\equiv& \beta \left[ U(x;\lambda) + p \, V(x) + \sum_{i=1}^N \mu_i \, N_i(x) + \cdots \right] . \label{equation:reduced-potential}
\end{eqnarray}
Here, the collection of thermodynamic and alchemical parameters $\theta \equiv \{\beta, \lambda, p, \mu, \ldots\}$ defines a \emph{thermodynamic state}.
The physical transformation of interest---such as a protein-ligand association process, or transfer free energy among phases---may involve a thermodynamic cycle that requires several free energy differences $\Delta f$ to be computed in order to produce an estimate of the overall desired quantity.

\todo[inline]{JDC: Should we give a few examples, along with illustrative figures?}

%%%%%%%%%%%%%%%%%%%%
%              Prerequesites                %
%%%%%%%%%%%%%%%%%%%%
\section{Prerequisites}
\label{sec:pre}
%Here you would identify prerequisites/background knowledge that are assumed by your work and your checklist which you view as critical, ideally giving links to good sources on these topics.
%Checklists are normally focused on errors made by users with training and experience in molecular simulations, so you can assume a basic familiarity with the fundamentals of molecular simulations.

In this review, we assume a basic familiarity with the principles of molecular mechanics, molecular dynamics simulations, statistical mechanics, and the biophysics of protein-ligand association.
This document focuses on aspects specific to alchemical free energy calculations, which remain an advanced topic within the molecular simulations area.
Thus, if you are a beginner in the area of molecular simulations, not all background you need will be covered within the scope of this document.
For example, we assume that you are able to set up and conduct a successful equilibrium molecular dynamics simulation of a particular thermodynamic state of a system interest as a prerequisite for this document.

\section{Scope and Goals}
\label{sec:scope}

This document focuses on preparation, execution, and analysis of alchemical free energy calculations, especially on aspects of the molecular simulations used here which are unique to alchemical calculations.
A particular focus is on calculation of transfer free energies (hydration free energies, partition coefficients, etc.) and binding free energies (absolute and relative).

Not within our scope are advanced ligand binding topics such as:

\begin{itemize}
\item Covalent inhibition
\item Association which is not 1:1
\item Endpoint free energy methods
\item PMF binding free energy methods
\end{itemize}

Additionally, the choice of force field (for protein, ligand, ions, cosolvents, or cofactors) is outside of the scope of this work; here, we focus on attempting to ensure successful alchemical free energy calculations given a particular choice of system and force field.

\textbf{Goals:}
\begin{itemize}
\item Checklist and background for new practitioners of relative and absolute alchemical free energy calculations: What should you pay attention to in setting up and running calculations in common codes and why.
\item Provide guidance to authors as to what should be reported about their protocols in a Methods section, either conforming to standard or reporting where they did not conform and why. Useful to reviewers as well.
\end{itemize}
\section{Checklist}
\label{sec:checklist}
\todo[inline, color={green!20}]{ASJSM: @Volunteer This needs to be revised and expanded, some initial thoughts were just thrown in.}
An attempt at identifying most important checklist items.

\section{Step 0 -- What can be expected from alchemical simulations?}
\label{sec:step0}
\todo[inline]{Huafeng Xu please edit and comment}
%\begin{itemize}
%\item What level of accuracy you can expect?
%\item What timescales and how many transformations can you address given available computational resources?
%\item Can you even hope to tackle the problem you are attempting?
%\end{itemize}

Current alchemical free energy calculations seem to achieve, in favorable cases, RMS errors around 1-2 kcal/mol depending on force field, system, and a variety of other factors such as simulation time, sampling method, and whether the calculations employed are absolute or relative. However, the domain of applicability is a significant concern [cite Sherborne, Cournia], especially for relative calculations, which typically require a bound structure of a closely
related ligand as a starting point. Additional factors such as slow protein or ligand rearrangements, uncertainties in ligand binding mode, or charged ligands can make these calculations far less reliable and more of a research effort.

@@

Before even planning free energy calculations to study binding to a
particular target, it is important to assess what is known about the
system and its timescales and its suitability for free energy
calculations, as well as the \emph{purpose} of the calculations and
the amount of available computer resources. In some cases, predicting
accurate binding free energies for a particular target might be a
\emph{more} challenging effort than simply measuring them! This is
often the case when dealing with database screening problems, where
compounds might be easily and quickly available commercially for
testing and free energy calculations could consume a much larger
amount of resources. Free energy calculations thus typically only
appeal when (slow or costly) synthesis would be required to do
experiments or experiments are otherwise cost-prohibitive.

@@

Additionally, the requisite level of accuracy is important -- if the
goal is to guide lead optimization when many compounds will be
synthesized, free energy calculations can be appealing even with
accuracies in the 1-2 kcal/mol range [cite Brown/Mobley/Shirts,
  Klimovich/Mobley Perspective], but if the number of compounds to be
synthesized is very small, this accuracy may not be enough to provide
much value.

@@

For illustrative purposes, we assume that the actual changes in the
binding free energies for a set of new molecules also follow normal
distribution, that the standard deviation in the changes is $RT\ln 5$
(corresponding to a 5-fold change in the binding affinities), and that
1 in 10 new molecules have increased binding affinity ($\Delta\Delta G
\leq 0$).  Under such assumptions, the conditional probability in
Eq.~\ref{eqn:true-positive} can be easily computed.  If the accuracy
of free energy calculations is $\sigma = 1$kcal/mol, $P(\Delta\Delta
G<0|\Delta\Delta G^\dagger<0) = 0.35$, which means that out of every
10 molecules selected for predicted favorable free energy change, on
average 3.5 molecules will have actual favorable free energy change.
In other words, selection by free energy calculations yields 3.5 times
more molecules of improved affinities than selection without free
energy calculations.

@@

This analysis should be done up front, part of ``counting the cost''
of involvement in a particular project. In some cases, the result of
this analysis may be a conclusion that free energy calculations will
certainly not be feasible for the proposed problem.

An additional consideration is how much is known about your particular
target, ligand binding modes in the target, and any relevant motions
-- essentially, has it been studied enough to know whether it might be
suitable for free energy calculations? If the conclusion it is hardly
been studied, this is important to know going in as well as, if
initial calculations perform poorly, the effort may turn into an
attempt to understand the relevant sampling, force field, or system
preparation problems.

Thus, Step 0 is essentially about initially counting the cost of your
project: Can you even hope to tackle the problem you are thinking
about attempting given available resources and, if successful, will it
be worth it?

If you don't know whether your project is feasible, one option may be
to conduct a short exploratory study to assess whether calculations
seem feasible at reasonable computational cost for just a very small
number of ligands. Sometimes this can be sufficient to get an initial
idea of feasibility and how accurate the calculations might be for the
proposed target.

\section{Step 1 -- Simulation prerequisites}
\label{sec:step1}
\todo[inline]{ASJSM: review first pass from JM}

In principle in the limit of sufficient conformational sampling, the free energy changes estimated from an alchemical free energy calculation should be independent of the system's initial coordinates. However in practice because simulations of thermodynamic states are of finite duration (typically currently 1-100 ns per state) this is only true for certain classes of alchemical free energy calculations such as relative or absolute free energies of hydration of small and relatively rigid organic molecules. Host-guest or protein-ligand complexes typically exhibit slowly relaxing degrees of freedom that exceed significantly the duration of an alchemical free energy calculation. It is therefore generally important to carefully select input coordinates to obtain satisfactory results. 
For binding free energy calculations the ligand should normally be positioned in its putative binding site in an energetically reasonable conformation that entails no steric or electrostatic mismatch with the receptor. Attention should be paid to binding site water molecules that may form water mediated protein-ligand interactions, particularly in situations where exchange with bulk water may be slow on the simulation timescales. This happens typically in buried binding site. In some implementation water sampling may be facilitated by the use of Grand Canonical Monte Carlo methods, but finite sampling limitations may prevent sampling of the correct solvent distributions. Information about water positions may be obtained from inspection of experimental data, or computed using a variety of molecular modelling methodologies (REFS). 
If the binding site location is uncertain but the structure of the receptor is well defined and plausible binding sites are apparent, absolute free energy calculations may be used to compute the standard free energy of binding of the ligand to a set of binding sites. This requires the user to prepare input files describing the bound conformation in different putative binding sites (REF). The apparent binding free energy of the ligand may be obtained by combining the individual binding site free energies, which also indicate where the ligand is more likely to bind.
If the putative binding sites are not apparent because of for instance significant induced-fit effects it may be challenging to obtain meaningful free energies of binding. One would have to account for the free energy change for forming a binding site in the target receptor. However it may still be possible to obtain useful binding free energy estimates that may be compared between different ligands.

For relative binding free energy calculations of congeneric series it is rare that experimental data is available to define the bioactive conformation of each analogue. A common assumption is that the binding mode is conserved within a congeneric series and the whole series can be modelled on the basis of a reference structure taken from experimental data or obtained by preliminary calculations. While conservation of binding modes within congeneric series seems to be the case (REFS?), exception exists(REF), and weaker binding compounds are more likely to adopt diverse binding modes (REFS). There may also be ambiguity about the relative orientation of substituents, a typical example are ortho and meta substituted 6-membered aromatic rings. While both positions are formally equivalent, a 180 degree flip of the ring may not occur sufficiently frequently during simulations. This situation may be addressed by explicitly modelling different binding modes of the same ligand and combining later computed free energy differences for different binding modes into a relative free energies of binding.

@@

\iffalse
\begin{enumerate}
\item Generate geometry of initial state: Reference biomolecular simulation preparation best practices
\item Relative: Generate geometry of final state (Mey)
\begin{itemize}
\item e.g. ligand ideally should not clash with receptor, etc.; satisfy constraints that might be imposed by protocol such as overlapping atoms, etc.
\item reference biomolecular preparation setup practices for placing ligand into binding site, but elaborate on constraints that must be considered for relative free energy calculations
\end{itemize}
\end{enumerate}
\fi

\section{Step 2 -- Simulation protocol selection}
\label{sec:step2}

\subsection{Absolute and relative free energy calculations have some differences}

@@

Sampling all binding modes within a single free energy calculation is usually impractical without some form of enhanced sampling or at least Hamiltonian replica exchange~\cite{Wang:2013:JournalofComputer-AidedMolecularDesign} because barriers for binding mode interconversion result in kinetics which are too slow compared to simulation timescales~\cite{Mobley:2006:TheJournalofChemicalPhysics, Palma:2012:J.Comput.Chem., Mobley:2012:TheJournalofChemicalPhysics, Gill:2018:J.Phys.Chem.B}.
Hamiltonian exchange, coupled with appropriate restraints, can allow the ligand to relatively rapidly exchange between potential binding modes when non-interacting, accelerating sampling of binding modes~\cite{Wang:2013:JournalofComputer-AidedMolecularDesign}.
\todo[inline, color={yellow!40}]{DLM probably need more background refs on hamiltonian lambda exchange here.}

@@

\section{Step 3 -- Overview of available analysis techniques}
\label{sec:step3}
\todo[inline, color={red!40}]{JDC, BA: uncertainty estimation section}

@@

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% ARTICLE SETUP
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\title{Best Practices for Alchemical Free Energy Calculations: v\versionnumber}
\author[7]{Bryce K. Allen}
\author[1*]{John D. Chodera}
\author[2*]{Antonia S. J. S. Mey}
\author[2*]{Jenke Scheen}
\author[2*]{Julien Michel}
\author[3*]{David L. Mobley}
\author[1]{Levi N. Naden}
\author[4*]{Samarjeet Prasad}
\author[5*]{Julia E. Rice}
\author[1,9]{Andrea Rizzi}
\author[6*]{Michael Shirts}
\author[10*]{Huafeng Xu}

@@

\begin{abstract}
Alchemical free energy calculations can be a useful tool for predicting free energy differences associated with the transfer of small molecules from one environment to another.
The hallmark of these methods is the use of modified potential energy functions to represent \emph{alchemical} intermediate states that cannot exist in chemistry; by analyzing simulation data collected from a series of bridging alchemical thermodynamic states, transfer free energies (or differences in transfer free energies) can be computed with orders of magnitude less simulation time than observing the process spontaneously. 
While these methods are highly flexible, care must be taken in avoiding common pitfalls to ensure that computed free energy differences can be robust and reproducible for the chosen forcefield, and that appropriate corrections are included to permit comparison with experimental data.
In this paper, we review current best practices for several popular application domains of alchemical free energy calculations, including relative and absolute small molecule binding free energy calculations to biomolecular targets.
\todo[inline, color=green!20]{JDC: What about biopolymer mutations? -- ASJSM: I think this is out of scope}
\end{abstract}

@@

%%%%%%%%%%%%%%%%%%%%
%  Introduction    %
%%%%%%%%%%%%%%%%%%%%
\section*{What are alchemical free energy methods?}
\label{sec:intro}
Alchemical free energy calculations have become a mature technology for computing various properties related to the transfer of chemical species from one environment to another.
The domain of applicability for these calculations now involves such varied applications as the computation of protein-ligand binding free energies~\cite{binding-free-energies}, ligand selectivities~\cite{selectivity}, partition and distribution coefficients between different liquid phases (such as octanol-water partition coefficients~\cite{octanol-water-partition}), loss of affinity due to resistance mutations~\cite{hauser2018predicting,aldeghi2018accurate}, and changes in protein thermostability due to engineered mutations~\cite{seeliger2010protein,gapsys2016insights,gapsys2016accurate,aldeghi2019accurate}.

\begin{figure}
    \includegraphics[width=0.95\linewidth]{paper/figures/Fig1_dummy.pdf}
    \caption{What is alchemistry}
    \label{fig:fig1}
\end{figure}

\todo[inline, color={green!20}]{ASJSM: This section should be expanded a little, but in a very general way, any volunteers?}

%%%%%%%%%%%%%%%%%%%%
% Prerequesites    %
%%%%%%%%%%%%%%%%%%%%
\section*{Prerequisites and Scope}
\label{sec:pre}
%Here you would identify prerequisites/background knowledge that are assumed by your work and your checklist which you view as critical, ideally giving links to good sources on these topics.
%Checklists are normally focused on errors made by users with training and experience in molecular simulations, so you can assume a basic familiarity with the fundamentals of molecular simulations.

Is my problem suitable for an alchemical free energy calculation? How do I chose the right free energy protocol and run it? How do I analyze my simulations accurately? What software tools are available to simulate alchemical free energy protocols? These are all questions this best practices guide aims to answer. It is meant to serve as a comprehensive starting point for new practitioners, as well as a reference guide with a convenient checklist~\ref{sec:checklist} to help with standard alchemical simulation and analysis practices. We will give an overview of different software that is available and its capabilities. The checklist we provide can serve as a good guide for a methods section both to the author of a free energy paper as well as a reviewer.

For ease of reading, we assume a basic familiarity with the principles of molecular mechanics, molecular dynamics simulations, statistical mechanics, and the biophysics of protein-ligand association.If you are unfamiliar with these concepts and have never setup and run an equilibrium molecular dynamics simulation, we would recommend the following texts for getting started with these topics first. (* indicate non essential reading/tutorial for more depth and understanding):
\begin{itemize}
    \item Statistical mechanics and thermodynamics background:
    \item Understanding theory of how molecular simulations work~\cite{something else awesome}
    \item Docking and generating valid molecular structures~\cite{something awesome}
    \item Getting started with setting up and running equilibrium molecular dynamics simulations: Gromacs tutorial~\cite{gromacs}, Amber tutorial~\cite{amber}, [any other?]
\end{itemize}

Some of the theoretical background can seem daunting therefore we provide an essentials guide to the theory behind alchemical fee energy calculations in section~\ref{sec:theory}. We will cover topics that are essential to the  preparation, execution, and analysis of alchemical free energy calculations. Particular focus will be given to aspects of the molecular simulations which are unique to alchemical calculations, these include calculation of transfer free energies (hydration free energies, partition coefficients, etc.) and binding free energies (absolute and relative).
While we try to address as many methods and practices as possible the field of free energy calculations is broad and there are many advanced topics that would do justice to a best practices document of their own right. Below we provide a none exhaustive lists of topics we have not addressed with some references to provide starting points on these more advanced topics:

\begin{itemize}
\item How to deal with covalently bound inhibitors~\cite{awesome reference}
\item Cases where associations are not based on a $1:1$ ratio~\cite{awesome reference}. 
\item Endpoint free energy methods~\cite{again some awesomeness}
\item Free energy methods that use potential mean force~\cite{all the cityness}
\item Forcefield dependence for protein, ligand, ions, co-solvents, and co-factors. Different studies have looked at the influence of force fields and we assume you made an adequate choice for they system you would like to study.~\cite{all the force field papers} 
\end{itemize}

%%%%%%%%%%%%%%%%%%%%
% Theory basics    %
%%%%%%%%%%%%%%%%%%%%

\section*{Theoretical background}
\label{sec:theory}
\todo[inline, color={green!20}]{Info box what is free energy}
\todo[inline, color={green!20}]{Info box what is a thermodynamic cycle}

$\left[
      \begin{tabular}{@{\quad}m{0.8\columnwidth}@{\quad}}
          \raggedright%
          \textbf{What is a thermodynamic cycle?} \par
            There's a package called \texttt{url} designed for typesetting web addresses.
            Write \mbox{\texttt{\string\usepackage\string{url\string}}} in your preamble; this will provide the
            command \mbox{\texttt{\string\url}}. This command takes an address for the argument and
            will print it out with typewriter font. Furthermore, it is able to handle
            special characters in addresses like underscores and percent signs. It even
            enables hyphenation in addresses, which is useful for websites with a very
            long name.%
      \end{tabular}
\right]$
\newline
%May need to look at this again: https://tex.stackexchange.com/questions/66820/how-to-create-highlight-boxes-in-latex
The defining characteristic of alchemical free energy calculations is the use of a series of alchemically-modified potential functions $U(x; \lambda)$ in which an alchemical parameter $\lambda$ modulates interactions in a manner that cannot occur in real chemical systems.
One or more simulations are used to collect data from a multitude of alchemical states to compute a free energy difference between a chemical state ($\lambda_0$) and another chemical or alchemical reference state ($\lambda_1$),
\begin{eqnarray}
\Delta f &\equiv& f(\lambda_1) - f(\lambda_0) = - \ln \frac{Z(\lambda_1)}{Z(\lambda_0)} , \label{equation:dimensionless-free-energy-difference}
\end{eqnarray}
where the dimensionless free energy $f(\lambda) \equiv \beta F(\lambda)$ is given in terms of partition functions $Z(\lambda)$,
\begin{eqnarray}
Z(\lambda) &=& \int dx \, e^{-u(x; \lambda)} . \label{equation:partition-function-definition}
\end{eqnarray}
Here, the inverse thermal energy $\beta \equiv (k_B T)^{-1}$ where $k_B$ is the Boltzmann constant and $T$ is the absolute temperature, and the \emph{reduced potential} $u(x; \lambda)$~\cite{shirts2008statistically} is generally given by a trace over thermodynamic parameters with their conjugate dynamical variables,
\begin{eqnarray}
u(x;\lambda) &\equiv& \beta \left[ U(x;\lambda) + p \, V(x) + \sum_{i=1}^N \mu_i \, N_i(x) + \cdots \right] . \label{equation:reduced-potential}
\end{eqnarray}
Here, the collection of thermodynamic and alchemical parameters $\theta \equiv \{\beta, \lambda, p, \mu, \ldots\}$ defines a \emph{thermodynamic state}.
The physical transformation of interest---such as a protein-ligand association process, or transfer free energy among phases---may involve a thermodynamic cycle that requires several free energy differences $\Delta f$ to be computed in order to produce an estimate of the overall desired quantity.

\begin{figure}
    \includegraphics[width=0.95\linewidth]{paper/figures/Fig2_dummy.pdf}
    \caption{Thermodynamic cycles}
    \label{fig:fig1}
\end{figure}
\todo[inline, color=green!20]{This secion needs to be expanded and give a good introduction to the theoretical background without going into too much detail.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Different steps required in planning and running a free energy calculation   %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{What can be expected from alchemical simulations? -- Step 0}
\label{sec:step0}
\todo[inline, color={green!20}]{@everyone please review this section}
Step 0 is essentially about initially counting the cost of your
project: Can you even hope to tackle the problem you are thinking
about attempting given available resources and, if successful, will it
be worth it?
\subsection*{How accurate are current alchemical free energy calculations?}
Current alchemical free energy calculations seem to achieve, in favorable cases, RMS errors around 1-2 kcal/mol depending on force field, system, and a variety of other factors such as simulation time, sampling method, and whether the calculations employed are absolute or relative. However, the domain of applicability is a significant concern~\cite{Sherborne2016, Cournia2017}, especially for relative calculations, which typically require a bound structure of a closely
related ligand as a starting point. Additional factors such as slow protein or ligand rearrangements, uncertainties in ligand binding mode, or charged ligands can make these calculations far less reliable and more of a research effort.

@@

\subsection{Is the system of study suitable for alchemical free energy calculations?}
Before even planning free energy calculations to study binding to a
particular target, it is important to assess what is known about the
system and its timescales and its suitability for free energy
calculations, as well as the \emph{purpose} of the calculations and
the amount of available computer resources. In some cases, predicting
accurate binding free energies for a particular target might be a
\emph{more} challenging effort than simply measuring them! This is
often the case when dealing with database screening problems, where
compounds might be easily and quickly available commercially for
testing and free energy calculations could consume a much larger
amount of resources. Free energy calculations thus typically only
appeal when (slow or costly) synthesis would be required to do
experiments or experiments are otherwise cost-prohibitive.

@@

\subsection{Is the expected accuracy of the computation good enough?}
Additionally, the requisite level of accuracy is important -- if the
goal is to guide lead optimization when many compounds will be
synthesized, free energy calculations can be appealing even with
accuracies in the 1-2 kcal/mol range [cite Brown/Mobley/Shirts,
  Klimovich/Mobley Perspective], but if the number of compounds to be
synthesized is very small, this accuracy may not be enough to provide
much value.

@@

For illustrative purposes, we assume that the actual changes in the
binding free energies for a set of new molecules also follow normal
distribution, that the standard deviation in the changes is $RT\ln 5$
(corresponding to a 5-fold change in the binding affinities), and that
1 in 10 new molecules have increased binding affinity ($\Delta\Delta G
\leq 0$).  Under such assumptions, the conditional probability in
Eq.~\ref{eqn:true-positive} can be easily computed.  If the accuracy
of free energy calculations is $\sigma = 1$kcal/mol, $P(\Delta\Delta
G<0|\Delta\Delta G^\dagger<0) = 0.35$, which means that out of every
10 molecules selected for predicted favorable free energy change, on
average 3.5 molecules will have actual favorable free energy change.
In other words, selection by free energy calculations yields 3.5 times
more molecules of improved affinities than selection without free
energy calculations.
\todo[inline, color={green!20}]{@DM: A plot of this might be useful. I believe you have used/generated one of these before.}

@@

\subsection*{Can I afford the calculation?}

This analysis should be done up front, part of ``counting the cost''
of involvement in a particular project. In some cases, the result of
this analysis may be a conclusion that free energy calculations will
certainly not be feasible for the proposed problem.
\todo[inline, color={green!20}]{@everyone: Might be nice to do an example calculation X perturbation on GTX 1080s on a AWS cluster cost \$\$X. Synthesis costs \$\$Y. }

\subsection*{Is an explorative study what I want?}
An additional consideration is how much is known about your particular
target, ligand binding modes in the target, and any relevant motions
-- essentially, has it been studied enough to know whether it might be
suitable for free energy calculations? If the conclusion it is hardly
been studied, this is important to know going in as well as, if
initial calculations perform poorly, the effort may turn into an
attempt to understand the relevant sampling, force field, or system
preparation problems.

If you are unsure whether your project is feasible, one option may be
to conduct a short exploratory study to assess whether calculations
seem feasible at reasonable computational cost for just a very small
number of ligands. Sometimes this can be sufficient to get an initial
idea of feasibility and how accurate the calculations might be for the
proposed target.

\section{Simulation prerequisites -- Step 1}
\label{sec:step1}
\todo[inline,color={green!20}]{@everyone: please review}
The alchemical free energy protocols, which are discussed in section~\ref{sec:step2}, which can be used are tied in with what type of free energy you might want to compute, i.e. a free energy of binding or a free energy of hydration. Depending on the type of simulations different considerations have to be made for ligands, solvent, and host molecules (in the case of the estimation of free energies of binding).
\subsection*{Free energies of binding}
In principle in the limit of sufficient conformational sampling, the free energy changes estimated from an alchemical free energy calculation should be independent of the system's initial coordinates. However, in practice because simulations of thermodynamic states are of finite duration (typically currently 1-100 ns per state) this is only true for certain classes of alchemical free energy calculations such as relative or absolute free energies of hydration of small and relatively rigid organic molecules. Host-guest or protein-ligand complexes typically exhibit slowly relaxing degrees of freedom that exceed significantly the duration of an alchemical free energy calculation. It is therefore generally important to carefully select input coordinates to obtain satisfactory results. 
You might want to ask yourself the following questions before diving into your simulation setup. 
\begin{itemize}
    \item Do I have one or multiple good host structures? (e.g. a good resolution X-ray crystal of the protein target)
    \item Should I include burried waters, or other small molecules that can be found in an X-ray structure.
    \item Are my ligands part of a congeneric series? (simple R group substitutions around the same scaffold)
    \item Do I have information on one or all of the ligands binding sites (e.g. a X-ray structure)
\end{itemize}

As for any simulation, good care should be taken in selecting protein structures by looking at available X-ray structures in the protein database and potentially selecting multiple structures initially, to account for variability in receptor conformations as well as accuracy of available X-ray structures. Typically, clustering of receptor structures can be used to identify different receptor conformations near the binding site, as well as assessing relevant side chain placements from a the x-ray structure, see for example~\cite{mey etal, the other D3R paper I can't remember}. In terms of set up and other choices following general best practice guidelines is advisable~\cite{Braun2018}.

For binding free energy calculations using a congeneric set of ligands and therefore likely to select a relative free energy protocol (see~\ref{sec:step2}), other than the choice of forcefield for the parametrization of the ligands, some care has to be taken selecting binding poses for these ligands. Generally a common assumption for congeneric is that the binding mode is conserved within the series. Therefore if an X-ray structure of one of the ligands is available, this should be used to position the congeneric series in its putative binding site in an energetically reasonable conformation that entails no steric or electrostatic mismatch with the receptor. Get in the habit of looking at X-ray structures and their electron densities properly, as often the resolution of part of the ligand may be more based on the interpretation of the crystallographer than the available electron density. For example, looking at a cyclo-hexane ring density, a chair conformation is vastly more likely than that of a boat~\cite{maybe we can find some reference here?} and may just be poorly assigned.

Generally binding modes within congeneric series are conserved (REFS?), however, exception exists(REF), and weaker binding compounds are more likely to adopt diverse binding modes (REFS). There may also be ambiguity about the relative orientation of substituents, a typical example are ortho- and meta-substituted 6-membered aromatic rings. While both positions are formally equivalent, a 180 degree flip of the ring may not occur sufficiently frequently during simulations. Another scenario may be equatorial and axial subsituted saturated rings (e.g. cyclohexane derivatives). This situation may be addressed by explicitly modelling different binding modes of the same ligand and combining later computed free energy differences for different binding modes into a relative free energies of binding.

Congeneric series can contain sterioismoerms or enantiomers. Making sure that you model the correct enantiomers or sterioisomers is important.
\todo[inline, color=green!20]{Question: What is best practice here? I tend to pick one enantiomer and just model that, but experimentally they may have used a racemic mixture?}

Another aspect to look out for are binding site water molecules that may form water mediated protein-ligand interactions, particularly in situations where exchange with bulk water may be slow on the simulation timescales. This happens typically in buried binding site. If multiple protein X-ray structures are available, retaining conserved water molecules is an easy way to not miss waters. In cases where water molecules are known to play an important role in the binding, software implementations that use water sampling facilitated by Grand Canonical Monte Carlo methods may be a good way to go. For examples see \textcolor{red}{system1, system2, and system3}~\cite{x,y,z}.

How to align congeneric series:
\todo[inline, color=green!20]{Question: How do people align congeneric ligand series?}

@@

\begin{figure}
    \includegraphics[width=0.95\linewidth]{paper/figures/Fig2_dummy.pdf}
    \caption{a) Bad alignment of congeneric series b) good alignment}
    \label{fig:fig1}
\end{figure}

If the binding site location is uncertain but the structure of the receptor is well defined and plausible binding sites are identified, you are more likely to chose an absolute free energy protocol to compute the standard free energy of binding of the ligand to a set of binding sites. This requires the user to prepare input files describing the bound conformation in different putative binding sites (REF). The apparent binding free energy of the ligand may be obtained by combining the individual binding site free energies, which also indicate where the ligand is more likely to bind. In this case using a docking program to generate initial structures is the way to go. Different commercial and none commercial tools are available, such as: rDock, Vina, glide, or Flare to name a few~\cite{rDock, Vina, Glide, Flare}. 
If the putative binding sites are not apparent because of for instance significant induced-fit effects it may be challenging to obtain meaningful free energies of binding. One would have to account for the free energy change for forming a binding site in the target receptor. However, it may still be possible to obtain useful binding free energy estimates that may be compared between different ligands.

\subsection*{Free energies of hydration or partition coefficients}
Preliminary considerations necessary for using free energy methods to compute partition coefficients are much generally much more straight forward. Using even something such as bable to generate a 3D minimised structure of the solute you would like to compute a free energy of hydration or partition coefficient can be sufficient for preparation of these types of simulations. However, in these cases a careful choice of forcefields, as well as water models or organic solvents is essential. See for example XX.et al~\cite{xxx} for a good discussion on these choices.

\section*{What simulation protocol should I choose? -- Step 2}
\label{sec:step2}
\todo[inline, color=green!20]{Just a note as a reminder: Chirality!}
\subsection{Absolute and relative free energy calculations have some differences}

@@

Sampling all binding modes within a single free energy calculation is usually impractical without some form of enhanced sampling or at least Hamiltonian replica exchange~\cite{Wang:2013:JournalofComputer-AidedMolecularDesign} because barriers for binding mode interconversion result in kinetics which are too slow compared to simulation timescales~\cite{Mobley:2006:TheJournalofChemicalPhysics, Palma:2012:J.Comput.Chem., Mobley:2012:TheJournalofChemicalPhysics, Gill:2018:J.Phys.Chem.B}.
Hamiltonian exchange, coupled with appropriate restraints, can allow the ligand to relatively rapidly exchange between potential binding modes when non-interacting, accelerating sampling of binding modes~\cite{Wang:2013:JournalofComputer-AidedMolecularDesign}.
\todo[inline, color={yellow!40}]{DLM probably need more background refs on Hamiltonian lambda exchange here.}

@@

\section{Step 3 -- Overview of available analysis techniques}
\label{sec:step3}
\todo[inline, color={red!40}]{JDC, BA: uncertainty estimation section}
\begin{enumerate}

@@

\section{Checklist}
\label{sec:checklist}
\todo[inline, color={green!20}]{ASJSM: @Volunteer This needs to be revised and expanded, some initial thoughts were just thrown in.}
An attempt at identifying most important checklist items.
